---
title: "covid 19"
author: "Desmond Koutso"
date: "Summer 2022"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex

$H_O$: Dying is not dependent of sex.  

$H_A$: Dying is dependent of sex.  


## Methods

Both Sex and result variables are categorical, so it is a cat~cat.  
the chi-squared and the fisher exact test will be used.  



Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

From this table, more females are alive then males, and the number of dead males is greater than that of dead females.    

## We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this

```{r}
dd2 <- kRdata %>% group_by(Sex,result) %>% summarize(count=n()) %>%
mutate(prcnt=count/sum(count))
```


# the group_by followed by summarize(count=n()) 

```{r}
basicC <- ggplot(dd2,aes(x=Sex,y=count,fill=result))
basicC + geom_bar(stat="identity",position="dodge")
```

Based on this plot, the frequency of female alive is greater than that of the males alive, and the frequency of dead males is greater than that of females alive.      
  
# Now for percentage plot

```{r}
basicCC <- ggplot(dd2,aes(x=Sex,y=prcnt*100,fill=result))
basicCC + geom_bar(stat="identity", position = "dodge")
```

This chart shows that men survive covid19 nearly as often as women do. Even though it still shows
that males die more often then females, but it isn't by much.    
  
## Numerical results


```{r}
table1 <- xtabs(~Sex + result, data=kRdata)
rowPerc(table1)
colPerc(table1)
```
The table shows that females survive covid19 98.7% of the time, while men survive covid19 97% of
the time. These numbers are very close for both males and females. Females die 1.4% of the time
and males die 3% of the time when they contract covid19.  
  

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
#fisherXY(kRdata$Sex, kRdata$result)
```

```{r}
fisher.test(table(kRdata$Sex, kRdata$result))
```

The chi-square test above shows our p-value is 0.0012 which is less than 0.05. Since it is less
than the 0.05 we reject the null hypothesis. The fisher test gives us a p-value of 0.0000000004
and an odds ratio of 2.3 making men 2.3 times more likely to die compared to females.    

## Conclusions

We investigated whether men and women are equally likely to die of covid19 from data collected in
South Korea through May 2019. $H_O$: Dying is not dependent on sex and $H_A$: Dying is dependent
on sex. We have a p-value of 0.0012 from our chi-squared test and therefore we reject the null
hypothesis. Lastly we looked at the fisher test and it's p-value was 0.0000000004 and the odds
ration of 2.3.  
