---
title: "covid 19"
author: "Desmond Koutso"
date: "Summer 2022"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex

$H_O$: Dying is not dependent of sex.  

$H_A$: Dying is dependent of sex.  


## Methods

Both Sex and result variables are categorical, so it is a cat~cat.  
the chi-squared and the fisher exact test will be used.  



Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

From this table, more females are alive then males, and the number of dead males is greater than that of dead females.    

## We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this


```{r}
barchartGC(~result + Sex,data=kRdata)
barchartGC(~result + Sex,data=kRdata, type="percent")
```

  Based off of the first bar chart, men seem to be more susceptible to Coronacirus than women. Men have a higher death rate compared to women. Women have a higher alive rate compared to men.  Approximately, the frequency for men living is 750 lower than that of women. Based off of the second bar chart, women have a 55% survival rate while men have a 40% survival rate. Women have a higher survival rate by 15%. The death percentage for men is higher, as expected. Men death rate is at 60% while women death rate is at 45%. Men death rate is higher by 15%. 


## Numerical  Descriptive results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

The data shows that 36 women  out of 100 people died out in the dead group. While 64 men out of 100 people died out in the dead group. More men are dying than women. Off of the percentages, women have a 98.7% survival rate while men ave a 97% survival rate. Women have a higher survival rate by 1.7%. Men have a higher death rate by 1.6% because 3 - 1.4 = 1.6. 


## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The null hypothesis has to be rejected because the p-value is .001. The p-value is not greater than .05 so the likelihood of getting these exact results is way too low. The fisher exact table is also significant in showing that we reject the null hypothesis because the p-value is .0009. Both of our p-values for the chi-square test and the fisher exact test are lower than .05, so we reject the null hypotheses. The sample estimate states that our odds ratio is 2.3 to 1. Therefore, the odds are 2.3 to 1 that death is related to the two sexes differently. That there is a 2.3 to 1 odds ratio that there will end up being a difference between the death rate of the two sexes. Death and sex are dependent but results are different for male & female. We can reject the null hypothesis. 


## Conclusions

The original question was if death was independent from sex. We can conclude that death isn't independent from sex. Dying is dependent on the sex. Our two bar charts, numerical data, inferential data, chi squared test, and the fisher's exact test all provide evidence towards my claim. Both p-values from the chi-squared test and the fisher's exact test are lower than .05. Thus, we have enough proof to reject the null hypotheses. Men have a higher likelihood of dying from Coronavirus compared to women. Women have a higher survival rate for Coronavirus compared to men. Those are my closing statements, thank you. 

