---
title: "covid 19"
author: "Desmond Koutso"
date: "Summer 2022"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex

$H_O$: Dying is not dependent of sex.  

$H_A$: Dying is dependent of sex.  


## Methods

Both sex and Sex variables are categorical, so it is a cat~cat.  
the chi-squared and the fisher exact test will be used.  



Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

From this table, more females are alive then males, and the number of dead males is greater than that of dead females.  

## We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this


```{r}
barchartGC(~sex + Sex,data=kRdata)
barchartGC(~sex + Sex,data=kRdata, type="percent")

```

Based on this plot, the frequency of female is greater than that of the males.    
  
  
## Numerical results


```{r}
table1 <- xtabs(~sex + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```
From the numerical results, there is no difference as the number of male is equal to the number of females.   

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```


from the table above, the P-value is0.0000000000000002, so we reject the null hypothesis by saying that, dying is dependent of sex.  

## Conclusions

The question was if dying is depend of sex?  
so, we conclude that yes, dying is dependent of sex.
